---
title: "HeartAttack&Prediction Analysis"
author: "Malay Raj"
date: "2023-09-26"
output:
  word_document: default
  html_document: default
---

## HEART ATTACK & PREDICTION ANALYSIS:

This dataset contains information about people's health, including details such as age, sex, chest pain, etc.

## Loading Required Packages.

```{r}
library(tidyverse)
library(janitor)
library(dplyr)
library(ggplot2)
library(skimr)
library(plyr)
library(lubridate)
library(plotly)
library(gridExtra)
```

## Displaying my current working directory.

```{r}
getwd()
```

## Collecting the data.

```{r}
heart = read.csv("C:/Heart Attack & Prediction Analysis/heart.csv")
```

## Displaying the first six rows of dataset.

```{r}
head(heart)
```

## Checking the dimension of the dataset.

```{r}
dim(heart)
```

## Displaying the column names of our dataframe.

```{r}
colnames(heart)
```

## Inspecting the dataframe and look for inconguencies.

```{r}
str(heart)
```

## Chaning few names of few columns in our dataset.

```{r}
heart = plyr::rename(heart, replace=c("cp" = "chest_pain",
                                      "fbs" = "fasting_blood_sugar",
                                      "exng" = "exercise_induced_angina",
                                      "slp" = "slope",
                                      "thall" = "thalassemia",
                                      "output" = "target"))
```

## Checking the column names of our new columns.

```{r}
colnames(heart)
```

## Checking for null values in our dataset.

```{r}
null_sum <- colSums(is.na(heart))
null_sum
```

## Checking for any duplicate values in our dataset.

```{r}
anyDuplicated(heart)
```

## Removing those duplicate values.

```{r}
heart = unique(heart)
anyDuplicated(heart)
```

## Checking for the current dimensions of our dataset.

```{r}
dim(heart)
```

## Seeing how many observations fall under each of these columns.

```{r}
table(heart$age)
```

```{r}
table(heart$sex)
```

```{r}
table(heart$chest_pain)
```

```{r}
table(heart$trtbps)
```

```{r}
table(heart$chol)
```

```{r}
table(heart$fasting_blood_sugar)
```

```{r}
table(heart$restecg)
```

```{r}
table(heart$thalachh)
```

```{r}
table(heart$exercise_induced_angina)
```

```{r}
table(heart$oldpeak)
```

```{r}
table(heart$slope)
```

```{r}
table(heart$caa)
```

```{r}
table(heart$thalassemia)
```

```{r}
table(heart$target)
```

## CONDUCTNG DESCRIPTIVE ANALYSIS:

Descriptive analysis on few columns of our dataset.

```{r}
summary(heart$age)
```

```{r}
summary(heart$trtbps)
```

```{r}
summary(heart$chol)
```

```{r}
summary(heart$thalachh)
```

```{r}
summary(heart$oldpeak)
```

Comparing different columns with each other from each of our dataset.

```{r}
aggregate(heart$trtbps ~ heart$age, data = heart, FUN = mean)
aggregate(heart$trtbps ~ heart$age, data = heart, FUN = median)
aggregate(heart$trtbps ~ heart$age, data = heart, FUN = max)
aggregate(heart$trtbps ~ heart$age, data = heart, FUN = min)
```

```{r}
aggregate(heart$chol ~ heart$age, data = heart, FUN = mean)
aggregate(heart$chol ~ heart$age, data = heart, FUN = median)
aggregate(heart$chol ~ heart$age, data = heart, FUN = max)
aggregate(heart$chol ~ heart$age, data = heart, FUN = min)
```

```{r}
aggregate(heart$thalachh ~ heart$age, data = heart, FUN = mean)
aggregate(heart$thalachh ~ heart$age, data = heart, FUN = median)
aggregate(heart$thalachh ~ heart$age, data = heart, FUN = max)
aggregate(heart$thalachh ~ heart$age, data = heart, FUN = min)
```

```{r}
aggregate(heart$oldpeak ~ heart$age, data = heart, FUN = mean)
aggregate(heart$oldpeak ~ heart$age, data = heart, FUN = median)
aggregate(heart$oldpeak ~ heart$age, data = heart, FUN = max)
aggregate(heart$oldpeak ~ heart$age, data = heart, FUN = min)
```

```{r}
aggregate(heart$age ~ heart$sex, data = heart, FUN = mean)
aggregate(heart$age ~ heart$sex, data = heart, FUN = median)
aggregate(heart$age ~ heart$sex, data = heart, FUN = max)
aggregate(heart$age ~ heart$sex, data = heart, FUN = min)
```

```{r}
aggregate(heart$trtbps ~ heart$sex, data = heart, FUN = mean)
aggregate(heart$trtbps ~ heart$sex, data = heart, FUN = median)
aggregate(heart$trtbps ~ heart$sex, data = heart, FUN = max)
aggregate(heart$trtbps ~ heart$sex, data = heart, FUN = min)
```

```{r}
aggregate(heart$thalachh ~ heart$sex, data = heart, FUN = mean)
aggregate(heart$thalachh ~ heart$sex, data = heart, FUN = median)
aggregate(heart$thalachh ~ heart$sex, data = heart, FUN = max)
aggregate(heart$thalachh ~ heart$sex, data = heart, FUN = min)
```

```{r}
aggregate(heart$oldpeak ~ heart$sex, data = heart, FUN = mean)
aggregate(heart$oldpeak ~ heart$sex, data = heart, FUN = median)
aggregate(heart$oldpeak ~ heart$sex, data = heart, FUN = max)
aggregate(heart$oldpeak ~ heart$sex, data = heart, FUN = min)
```

## Comparing different columns of our dataframe with Age and Sex of people.

```{r}
aggregate(heart$chest_pain ~ heart$sex + heart$age, FUN = mean)
aggregate(heart$trtbps ~ heart$sex + heart$age, FUN = mean)
aggregate(heart$chol ~ heart$sex + heart$age, FUN = mean)
aggregate(heart$fasting_blood_sugar ~ heart$sex + heart$age, FUN = mean)
aggregate(heart$restecg ~ heart$sex + heart$age, FUN = mean)
aggregate(heart$thalachh ~ heart$sex + heart$age, FUN = mean)
aggregate(heart$exercise_induced_angina ~ heart$sex + heart$age, FUN = mean)
aggregate(heart$oldpeak ~ heart$sex + heart$age, FUN = mean)
aggregate(heart$slope ~ heart$sex + heart$age, FUN = mean)
aggregate(heart$caa ~ heart$sex + heart$age, FUN = mean)
aggregate(heart$thalassemia ~ heart$sex + heart$age, FUN = mean)
aggregate(heart$target ~ heart$sex + heart$age, FUN = mean)
```

## Changing the values of few columns which will help us in better under standing of the visualization.

```{r}
heart$sex <- ifelse(heart$sex == 0, "female", "male")
```

```{r}
heart$chest_pain <- ifelse(heart$chest_pain == 0, "typical angina",
                           ifelse(heart$chest_pain == 1, "atypical angina",
                                  ifelse(heart$chest_pain == 2, "non-anginal pain", 
                                         ifelse(heart$chest_pain == 3, "asymptomatic", NA))))
```

```{r}
heart$fasting_blood_sugar <- ifelse(heart$fasting_blood_sugar == 1, "true", "false")
```

```{r}
heart$restecg <- ifelse(heart$restecg == 0, "normal",
                        ifelse(heart$restecg == 1, "ST-T wave abnormality",
                               ifelse(heart$restecg == 2, "showing probable", NA)))
```

```{r}
heart$exercise_induced_angina <- ifelse(heart$exercise_induced_angina == 1, "yes", "no")
```

```{r}
heart$slope <- ifelse(heart$slope == 0, "unsloping",
                        ifelse(heart$slope == 1, "flat",
                               ifelse(heart$slope == 2, "downsloping", NA)))
```

```{r}
heart$thalassemia <- ifelse(heart$thalassemia == 0, "null",
                        ifelse(heart$thalassemia == 1, "fixed defect",
                               ifelse(heart$thalassemia == 2, "normal",
                                      ifelse(heart$thalassemia == 3, "reversable defect", NA))))
```

```{r}
heart$target <- ifelse(heart$target == 0, "less chance", "more chance")
```

## Data Visualization:

Let's see the count of total male and female in our dataset.

```{r}
ggplot(heart, aes(x = sex, fill = sex)) +
  geom_bar() +
  labs(title = "Count Of SEXs", x = "Genders", y = "Count") +
  theme_minimal()
```

Let's see the count of different types of chest pain in our dataset.

```{r}
ggplot(heart, aes(x = chest_pain, fill = chest_pain)) +
  geom_bar() +
  labs(title = "Count Of different types of chest_pains", x = "Chest_pain", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Let's see the percentage distribution of fasting blood sugars.

```{r}
data <- as.data.frame(table(heart$fasting_blood_sugar))
pie_chart <- ggplot(data, aes(x = "", y = Freq, fill = Var1)) +
  geom_bar(stat = "identity") +
  coord_polar(theta = "y") +
  geom_text(aes(label = paste0(round((Freq/sum(Freq))*100), "%")),
            position = position_stack(vjust = 0.5)) +
  labs(title = "Fasting Blood Sugar Distribution",
       fill = "Fasting Blood Sugar") +
  theme_void()
print(pie_chart)
```

Let's see the count of resting electrocardiographic results in our dataset.

```{r}
ggplot(heart, aes(x = restecg, fill = restecg)) +
  geom_bar() +
  labs(title = "Count of Resting Electrocardiographic Results", x = "Resting Electrocardiographic Results", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Let's see the percentage distribution of exercise induced angina in our dataset.

```{r}
data <- as.data.frame(table(heart$exercise_induced_angina))
pie_chart <- ggplot(data, aes(x = "", y = Freq, fill = Var1)) +
  geom_bar(stat = "identity") +
  coord_polar(theta = "y") +
  geom_text(aes(label = paste0(round((Freq/sum(Freq))*100), "%")),
            position = position_stack(vjust = 0.5)) +
  labs(title = "Exercise Induced Angina Distribution",
       fill = "Exercise Induced Angina") +
  theme_void()
print(pie_chart)
```

Let's see the count of slopes in our dataset.

```{r}
ggplot(heart, aes(x = slope, fill = slope)) +
  geom_bar() +
  labs(title = "Count of Slope", x = "Slope", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Let's see the distribution in pie chart of caa.

```{r}
data <- as.data.frame(table(heart$caa))
pie_chart <- ggplot(data, aes(x = "", y = Freq, fill = Var1)) +
  geom_bar(stat = "identity") +
  coord_polar(theta = "y") +
  labs(title = "caa",
       fill = "caa") +
  theme_void()
print(pie_chart)
```

Let's see the count of different thalassemia in our dataset.

```{r}
ggplot(heart, aes(x = thalassemia, fill = thalassemia)) +
  geom_bar() +
  labs(title = "Count of Thalassemia", x = "Thalassemia", y = "Count") +
  theme_minimal() +
  theme(legend.position = "top")
```

Let's see the percentage distribution of people having chances of heart attack.

```{r}
data <- as.data.frame(table(heart$target))
pie_chart <- ggplot(data, aes(x = "", y = Freq, fill = Var1)) +
  geom_bar(stat = "identity") +
  coord_polar(theta = "y") +
  geom_text(aes(label = paste0(round((Freq/sum(Freq))*100), "%")),
            position = position_stack(vjust = 0.5)) +
  labs(title = "target",
       fill = "target") +
  theme_void()
print(pie_chart)
```

Let's see the count of chest pains in different genders.

```{r}
sex_count <- heart %>%
  group_by(sex, chest_pain) %>%
  summarise(sex_count = n(), .groups = "drop")
ggplot(sex_count, aes(x = sex, y = sex_count, fill = chest_pain)) +
  geom_col(position = "dodge") +
  labs(title = "Count of SEXs by different Chest Pains",
       x = "Sex",
       y = "Count") +
  theme_minimal()
```

Let's see the count of fasting blood sugar in different genders.

```{r}
count_fbs <- heart %>%
  group_by(sex, fasting_blood_sugar) %>%
  summarise(count_fbs = n(), .groups = "drop")
ggplot(count_fbs, aes(x = sex, y = count_fbs, fill = fasting_blood_sugar)) +
  geom_col(position = "dodge") +
  labs(title = "Count of SEXs by Fasting Blood Sugar",
       x = "Sex",
       y = "Count") +
  theme_minimal() +
  theme(legend.position = "top")
```

Let's see the count of resting ecg results in different genders.

```{r}
count_restecg <- heart %>%
  group_by(sex, restecg) %>%
  summarise(count_restecg = n(), .groups = "drop")
ggplot(count_restecg, aes(x = sex, y = count_restecg, fill = restecg)) +
  geom_col(position = "dodge") +
  labs(title = "Count of SEXs by Resting Electrocardiographic Results",
       x = "Sex",
       y = "Count") +
  theme_minimal() +
  theme(legend.position = "top")
```

Let's see the count of exercise induced angina in different genders.

```{r}
count_eia <- heart %>%
  group_by(sex, exercise_induced_angina) %>%
  summarise(count_eia = n(), .groups = "drop")
ggplot(count_eia, aes(x = sex, y = count_eia, fill = exercise_induced_angina)) +
  geom_col(position = "dodge") +
  labs(title = "Count of SEXs by Exercise Induced Angina",
       x = "Sex",
       y = "Count") +
  theme_minimal() +
  theme(legend.position = "top")
```

Let's see the count of slopes in different genders.

```{r}
count_slope <- heart %>%
  group_by(sex, slope) %>%
  summarise(count_slope = n(), .groups = "drop")
ggplot(count_slope, aes(x = sex, y = count_slope, fill = slope)) +
  geom_col(position = "dodge") +
  labs(title = "Count of SEXs by Slope",
       x = "Sex",
       y = "Count") +
  theme_minimal() +
  theme(legend.position = "top")
```

Let's see the count of different thalassemias in different genders.

```{r}
count_thala <- heart %>%
  group_by(sex, thalassemia) %>%
  summarise(count_thala = n(), .groups = "drop")
ggplot(count_thala, aes(x = sex, y = count_thala, fill = thalassemia)) +
  geom_col(position = "dodge") +
  labs(title = "Count of SEXs by Thalassemia",
       x = "Sex",
       y = "Count") +
  theme_minimal() +
  theme(legend.position = "top")
```

Finally, let's see the chance of getting heart attack in different genders.

```{r}
count_target <- heart %>%
  group_by(sex, target) %>%
  summarise(count_target = n(), .groups = "drop")
ggplot(count_target, aes(x = sex, y = count_target, fill = target)) +
  geom_col(position = "dodge") +
  labs(title = "Count of SEXs by Target",
       x = "Sex",
       y = "Count") +
  theme_minimal() +
  theme(legend.position = "top")
```

